Can Cultural Transmission Explain The Evolution of Cooperative Behavior?

### 2020-03-05 ms/main.tex
small changes

\textbf{Kin selection}, proposing that natural selection can favor cooperative behavior between kin. The importance of relatedness to the evolution of cooperation and altruism was shown by~\citet{hamilton1964genetical}. According to Hamilton, kin selection causes allele to increase in frequency when the genetic relatedness $r$ of a recipient to an actor multiplied by the benefit $b$ to the recipient is greater than the reproductive cost $c$ to the actor. This is also known as Hamilton's Rule:

@@

All the above theories assume that cooperation is genetically determined. This raises the question, is it possible that cooperation is determined by non-genetic factors?
Recent work by \citet{lewin2017microbes} sheds some light on this question. 
\citet{lewin2017microbes} have hypothesised that microbes that manipulate their hosts to act altruistically can be favored by selection, and may play a role in the widespread occurrence of cooperative behavior. Indeed, it has been shown that microbes can mediate behavioral changes in their hosts~\citep{poulin2010parasite,dobson1988population}. Therefore, natural selection on microbes may favor manipulation of the host so that it cooperates with others. Microbes can be transmitted \emph{horizontally} from one host to another during host interactions. Following horizontal transfer, the recipient host may carry microbes that are closely related to the microbes of the donor host, even when the two hosts are (genetically) unrelated~\citep{lewin2017microbes}. Microbes can also transfer vertically, from parent to offspring. % yr: cite 10.1126/science.aat7164
As a result, a microbe that induces its host to cooperate with another host and thereby increases the other host fitness will  increase the vertical transmission of the microbes of the receiving individual. Kin selection among microbes could therefore favor microbes that induce cooperative behavior in their hosts, thereby increasing the transmission of their microbial kin.
%\\While the hypothesis that microbes could manipulate their hosts to act altruistically is mathematically feasible~\citep{lewin2017microbes}, there are still no empirical evidence of whether microbes indeed mediate altruistic behavior of their hosts.

% subsection
\subsection*{Cultural evolution of cooperation}
\citet{lewin2017microbes} have demonstrated that \emph{non-vertical} transmission can help to explain the evolution of cooperative behavior. 
Non-vertical transmission could be either a horizontal or oblique. Horizontal transmission occurs between individuals from the same generation. Oblique transmission occurs from an adult to an offspring. 
Evolution under either of these transmission models is can be be more rapid than under pure vertical transmission~\citep{ram2018evolution}. % (yr: not sure that this is a good reference)

@@

\paragraph*{Without oblique transmission, $v=1$.} 
With only vertical transmission eq.~\ref{eq:vertical_oblique_transmission} becomes
$\hat{p} =  \tilde{p}$,
and eq.~\ref{eq:nextgen_parents} for the change in frequency $p'$ of phenotype $A$ among parents can be written as

@@

\textbf{Result 1.}
If 
\begin{equation}\label{result1}
c \cdot (1-T_B) < b \cdot  \alpha T_A  + (T_A - T_B) [1 + b (p(1-\alpha) - 1)]
\end{equation}
then $\tilde{p}' > \tilde{p}$, and the frequency of phenotype $A$ among parents increases.

>>>>

\paragraph{Kin selection theory} suggests that natural selection can favor cooperative behavior between kin. The importance of relatedness to the evolution of cooperation and altruism was shown by~\citet{hamilton1964genetical}. According to Hamilton, kin selection causes allele to increase in frequency when the genetic relatedness $r$ of a recipient to an actor multiplied by the benefit $b$ to the recipient is greater than the reproductive cost $c$ to the actor. This is also known as Hamilton's Rule:

@@

\paragraph{}
All the above theories assume that cooperation is genetically determined. This raises the question, is it possible that cooperation is determined by non-genetic factors?
Recent work by \citet{lewin2017microbes} sheds some light on this question. 
\citet{lewin2017microbes} have hypothesised that microbes that manipulate their hosts to act altruistically can be favored by selection, and may play a role in the widespread occurrence of cooperative behavior. Indeed, it has been shown that microbes can mediate behavioral changes in their hosts~\citep{poulin2010parasite,dobson1988population}. Therefore, natural selection on microbes may favor manipulation of the host so that it cooperates with others. Microbes can be transmitted \emph{horizontally} from one host to another during host interactions. Following horizontal transfer, the recipient host may carry microbes that are closely related to the microbes of the donor host, even when the two hosts are (genetically) unrelated~\citep{lewin2017microbes}. Microbes can also transfer vertically, from parent to offspring. % yr: cite 10.1126/science.aat7164
As a result, a microbe that induces its host to cooperate with another host and thereby increases the other host fitness will  increase the vertical transmission of the microbes of the receiving individual. Kin selection among microbes could therefore favor microbes that induce cooperative behavior in their hosts, thereby increasing the transmission of their microbial kin.
%\\While the hypothesis that microbes could manipulate their hosts to act altruistically is mathematically feasible~\citep{lewin2017microbes}, there are still no empirical evidence of whether microbes indeed mediate altruistic behavior of their hosts.

% subsection
\subsection*{Cultural evolution of cooperation}
\citet{lewin2017microbes} have demonstrated that \emph{non-vertical transmission} can help to explain the evolution of cooperative behavior. 
Non-vertical transmission could be either a horizontal or oblique. Horizontal transmission occurs between individuals from the same generation. Oblique transmission occurs from an adult to an offspring. 
Evolution under either of these transmission models is can be be more rapid than under pure vertical transmission~\citep{ram2018evolution}. % (yr: not sure that this is a good reference)

@@

With only vertical transmission eq.~\ref{eq:vertical_oblique_transmission} becomes
$\hat{p} =  \tilde{p}$,
and eq.~\ref{eq:nextgen_parents} for the change in frequency $p'$ of phenotype $A$ among parents can be written as

@@

\textbf{Result 1: Vertical and horizontal transmission of cooperation.}
If 
\begin{equation}\label{result1}
c \cdot (1-T_B) < b \cdot  \alpha T_A  + (T_A - T_B) [1 + b (p(1-\alpha) - 1)]
\end{equation}
then $\tilde{p}' > \tilde{p}$, and the frequency of phenotype $A$ among parents increases.
